---
title: "5104 Homework 5"
author: "Ryan Christianson"
date: "September 26, 2018"
output: pdf_document
---

```{r, warning = FALSE, message = FALSE}
# global options
knitr::opts_chunk$set(comment = NA)

# libraries
library("ggplot2")
library("GGally")
library("data.table")
library("downloader")
library("fiftystater")
```

## Problem 1

Done.

## Problem 2

Done.

## Problem 3

They should be able to understand the true relationships in the data. For example uncover a dinosaur when plotting by a factor.

## Problem 4

```{r}
## part a - the mean function does this...
GetProp <- function(vec) {
  return(mean(vec))
}

## part b
set.seed(12345)
p4b.data <- matrix(rbinom(10, 1, prob = (30:40)/100), nrow = 10,
                   ncol = 10)

## part c
apply(p4b.data, 1, mean)
apply(p4b.data, 2, mean)

## part d
set.seed(12345)
p4d.data <- matrix(rbinom(100, 1, prob = (30:40)/100), nrow = 10,
                   ncol = 10)

apply(p4d.data, 1, mean)
apply(p4d.data, 2, mean)
```

c. It copied the same vector into each column of the data, so the row proportions are 1 or 0 and the column proportions are all the same.

## Problem 5

```{r, message = FALSE}
url5 <- "https://www2.isye.gatech.edu/~jeffwu/book/data/starch.dat"
p5.data <- read.table(url5, header = TRUE)
ggpairs(p5.data)

ggplot(data = p5.data, aes(x = thickness, y = strength, color = starch)) +
  geom_point() +
  labs(x = "Thickness", y = "Strength") +
  ggtitle("Distribution of Strength and Thickness Colored by Starch") +
  theme_bw()
```

I started off with a pairs plot to visualize the variables. Then I wanted to see a plot that included all variables, so I made a scatterplot colored by starch. I feel I can adequately see the structure of the data now.

## Problem 6

```{r}
# part a

# we are grabbing a SQL set from here
# http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip
# download the files, looks like it is a .zip
download("http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip",
         dest = "us_cities_states.zip")
# removed exdir = "./"
unzip("us_cities_states.zip")
# read in data, looks like sql dump, blah

# removed skip = 23 to read in the states that start with A
states <- fread(input = "./us_cities_and_states/states.sql",
                sep = "'", sep2 = ",",
                header = FALSE, select = c(2, 4))

colnames(states) <- c("State", "Code")
### YOU do the CITIES I suggest the cities_extended.sql
### may have everything you 
cities <- fread(input = "./us_cities_and_states/cities_extended.sql",
                sep = "'", sep2 = ",",
                header = FALSE, select = c(2, 4))
colnames(cities) <- c("City", "State")

# part b
cities.table <- table(cities$State)
cities.table
# don't use PR and DC because they are not in crime dataset
cities.table <- cities.table[!(names(cities.table) %in%  c("PR", "DC"))]
states <- states[!(states$Code %in% c("PR", "DC")), ]
num.cities <- data.frame(Cities = as.numeric(cities.table),
                         Code = names(cities.table))
states <- merge(states, num.cities, by = "Code")

# part c
CountLetter <- function(letter, state.name) {
  state.name <- tolower(state.name)
  return(sum(strsplit(state.name, "")[[1]] == letter))
}

letter.count <- data.frame(matrix(NA, nrow = 50, ncol = 26))
colnames(letter.count) <- letters
for (i in 1:50) {
  letter.count[i, ] <- sapply(letters, CountLetter,
                              state.name = states$State[i])
}

# part d
# https://cran.r-project.org/web/packages/fiftystater/vignettes/fiftystater.html

data("fifty_states") # this line is optional due to lazy data loading
crimes <- data.frame(state = tolower(rownames(USArrests)),
                     USArrests)

states$State <- tolower(states$State)
crimes <- merge(crimes, states, by.x = "state", by.y = "State")
# map_id creates the aesthetic mapping to the state name
# column in your data
ggplot(crimes, aes(map_id = state)) +
  geom_map(aes(fill = Cities), map = fifty_states) +
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", panel.background = element_blank())

states.3letter <- states[apply(letter.count, 1, max) > 3, ]
ggplot(states.3letter, aes(map_id = State)) +
  geom_map(aes(fill = Cities), map = fifty_states) +
    expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", panel.background = element_blank())
```

## Problem 7

Done.